### 2019-02-25 manuscript/methods.tex
editing through model comparisons in Results

Using the marginal likelihood values, we calculated thirteen different Bayes factors.
Six compared the models of polyploidy against each other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily. I find it useful to imagine Spencer reading this, for example.
% E: "Lumpable" is the math word for a particular relation among state codings, used in that Tarasov paper.  I don't think we want to get into defining it precisely.  Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.  Probably need to add to its caption.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
The number of species in each is shown under the corresponding circles in the top row.
These categories were then grouped in a manner appropriate to the states of each model.
For example, there are 34 species that are self-compatible and of unknown ploidy; these are coded as either $D$ or $P$ in the D/P models (uncertain, or consistent with either state), as $C$ in the I/C models, and as either $CD$ or $CP$ in the ID/CD/CP models.
In all cases, species were coded as either $A$ or $B$ in the hidden state models.
}
\label{figure:stateclassifications}
\end{figure}

\begin{table}
\begin{tabular}{@{}llcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Model& Model& Ploidy & Diploidization & Breeding  & Hidden & Parameters & Marginal \\
& Type & &  &System & State & & Log- Likelihood \\
1. D/P &BiSSE &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
2. D/P no $\delta$ & BiSSE  &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
3. D/P- A/B  & HiSSE &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
4. D/P no $\delta$-A/B &	HiSSE &	Yes & No &	No &	Yes &	10	&-1150.99\\
5. I/C &	BiSSE &	No &	 No	&Yes &	No &	 5 &  -1194.80 \\
6. I/C-A/B &	HiSSE &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
7. ID/P/CD & MuSSE &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
8. ID/P/CD no $\delta$ &	MuSSE &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
9. ID/P/CD-A/B & MuHiSSE &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
10.ID/P/CD no $\delta$-A/B &	 MuHiSSE	 & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
\end{tabular}
\caption{Marginal likelihoods for the ten models examined. States encoded with letter D: diploid, P: polyploid, C: self-compatible, I: self-incompatible, A: first state of hidden trait, B: second state of hidden trait, $\delta$: re-diploidization.}
\label{table:marginallike}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|c|c|c|}
\toprule
Ploidy Models & Breeding System Models & Ploidy and Breeding System Models \\ \midrule
{\begin{tabular}{lcccc}
 & 1 & 2 & 3 & 4 \\
1. D/P & $\cdot$	 &10.72 &	-37.24	&-31.94\\
2. D/P no $\delta$ &$\cdot$&$\cdot$ &	-47.97	&-42.66\\
\textbf{3. D/P- A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
4. D/P no $\delta$-A/B &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
\end{tabular}
}  & 
{\begin{tabular}{lcc}
 & 5 & 6\\
5. I/C &$\cdot$ & -39.43 \\
\textbf{6. I/C-A/B} &$\cdot$& $\cdot$ \\
& & \\
& & \\
\end{tabular}
} & 
{\begin{tabular}{lcccc}
& 7 & 8 & 9 & 10\\
7. ID/P/CD & $\cdot$	&1.36 & -44.15&-40.95 \\
8. ID/P/CD no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
\textbf{9. ID/P/CD-A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
\end{tabular}
}\\
\bottomrule
\end{tabular}
 \caption{Bayes factors in log scale. We compare every possible pair. Number models as indicated in  \cref{table:marginallike}. }
\label{table:bayesfactors}
\end{table}

>>>>

Using the marginal likelihood values, we calculated thirteen different Bayes factors.
Six compared the models of ploidy against one other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In mathematical terms, the D/P, I/C, and ID/CD/CP state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily. I find it useful to imagine Spencer reading this, for example.
% E: "Lumpable" is the math word for a particular relation among state codings, used in that Tarasov paper.  I don't think we want to get into defining it precisely.  Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.  I added to the caption and the text above.  Okay, or more work needed?
Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1]$.
There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
The number of species in each is shown under the corresponding circles in the top row. % FIXME: Rosana, what are the numbers for CP and ?P ?
These categories were then grouped in a manner appropriate to the states of each model.
For example, there are 34 species that are self-compatible and of unknown ploidy; these are coded as either $D$ or $P$ in the D/P models (uncertain, or consistent with either state), as $C$ in the I/C models, and as either $CD$ or $CP$ in the ID/CD/CP models.
In all cases, species were coded as either $A$ or $B$ in the hidden state models.
The models depicted in the different rows use state spaces that are not comparable with one another.
For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same and are not `lumpable' \citep{tarasov_2019}.
}
\label{figure:stateclassifications}
\end{figure}

\begin{table}
\begin{tabular}{lcccccc} \toprule
% \begin{tabular}{@{}lcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Model                        & Ploidy & Diploidization & Breeding  & Hidden & Num & Marginal \\
                             & &  &System & State & Parameters & Log-Likelihood \\ \midrule
1. D/P                       &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
2. D/P no $\delta$           &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
3. D/P+A/B                   &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
4. D/P+A/B no $\delta$       &	Yes & No &	No &	Yes &	10	&-1150.99\\
5. I/C                       &	No & No	&Yes &	No &	 5 &  -1194.80 \\
6. I/C+A/B                   &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
7. ID/CD/CP                  &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
8. ID/CD/CP no $\delta$      &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
9. ID/CD/CP+A/B              &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
10. ID/CD/CP+A/B no $\delta$ & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
\end{tabular}
\caption{
    The ten models and their marginal likelihoods.
    Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
    Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
\label{table:marginallike}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|c|c|c|}
\toprule
Ploidy Models & Breeding System Models & Ploidy and Breeding System Models \\ \midrule
{\begin{tabular}{lcccc}
 & 1 & 2 & 3 & 4 \\
1. D/P & $\cdot$	 &10.72 &	-37.24	&-31.94\\
2. D/P no $\delta$ &$\cdot$&$\cdot$ &	-47.97	&-42.66\\
\textbf{3. D/P+A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
4. D/P+A/B no $\delta$ &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
\end{tabular}
}  & 
{\begin{tabular}{lcc}
 & 5 & 6\\
5. I/C &$\cdot$ & -39.43 \\
\textbf{6. I/C+A/B} &$\cdot$& $\cdot$ \\
& & \\
& & \\
\end{tabular}
} & 
{\begin{tabular}{lcccc}
& 7 & 8 & 9 & 10\\
7. ID/CD/CP & $\cdot$	&1.36 & -44.15&-40.95 \\
8. ID/CD/CP no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
\textbf{9. ID/CD/CP+A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
\end{tabular}
}\\
\bottomrule
\end{tabular}
% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
\caption{
    Bayes factors for model comparisons.
    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}).
    Models are numbered in as \cref{table:marginallike}.
    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
    The best model in each set is written in bold.
    In each case, it is the most complex model of the set.
}
\label{table:bayesfactors}
\end{table}

### 2019-02-25 manuscript/results.tex
editing through model comparisons in Results

% E: I brutally shortened the text here and hence put all the models into one section for the "diversification" question.  But I think I retained all the ideas that were written out before.
% Much of what I removed was descriptions about which distributions are/aren't overlapping with one another.  I think that's easier for the reader to see in the figures than to read in the text.  But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?

@@

For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model, \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
% E: I removed the description of relative extinction results.  We don't discuss it for other models or show it in the main text, and I never really know what to make of that compound parameter.  But it's good that it's in the supp info.

For breeding system alone, we found a larger net diversification rate for SI and for SC species, in agreement with \citep{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
When a hidden state is added to the model (I/C+A/B), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
Thus, differences in net diversification are best explained by both breeding system and an unknown factor.
The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.

For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids (ID) was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
The association of ploidy with net diversification (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
%
When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is excluded (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.

@@

\subsection{Model selection}
% E: Maybe incorporate this into the previous subsection?  It could make more sense to organize by question rather than by statistical method (param estimates vs model comparisons).

In \cref{table:marginallike} we listed the marginal likelihood values calculated in log scale ($log(P(\mathbf{X}|M)$) for each of the models tested.
The different components included and excluded for each model as a a summary of the diagrams from \cref{figure:netdivall} are indicated in each model.
From the marginal likelihoods in log scale, the Bayes factors in log-scale($BF(M_0,M_1)=log(P(\mathbf{X}|M_0)-P(\mathbf{X}|M_0)$) were calculated as shown in table \cref{table:bayesfactors}.
After testing every single pair of polyploidy models (1-4) we found overwhelming evidence that the best polyploidy model is always the D/P+A/B, that is the model with hidden state and diploidization, that is Bayes factors were always greater than 2 when D/P+A/B was the first model in the Bayes factor equation or negative when D/P+A/B was the second model in Bayes factor equation.

For the two models following the evolution of breeding system, the I/C+A/B is the best choice between models 3 and 4 (BF(IC, IC+A/B)=-39.5<0).
%
The four models that follow the diversification linked to both polyploidy and breeding system (models 5-8) where compared in pairs.
Bayes factors show decisive evidence in favor of the IC/P/CD+A/B over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options \cref{table:bayesfactors}.
%
Therefore, the models that were chosen were always the ones containing the hidden trait, or diploidization parameter $\delta$ when ploidy state was considered.

@@

\begin{figure}
\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf}
  \caption{Net diversification rates for all models that include diploidization.}  
\label{figure:netdivall}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf}
  \caption{Net diversification rates for all models that do not include diploidization.}  
\label{figure:netdivnodip}
\end{figure}

@@

% E or R todo: which pathway from SI-D to SC-P is more common, based on rate estimates?  apply procedure from robertson2011 to rates from ID/CD/CP model

>>>>

% E: I brutally shortened the text here and hence put all the models into one section for the "diversification" question.  But I think I retained all the ideas that were written out before.
% Much of what I removed was descriptions about which distributions are/aren't overlapping with one another.  I think that's easier for the reader to see in the figures than to read in the text.  But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?
% For the former model comparison section, I moved the key conclusions into the paragraphs here and the mechanical details (bayes factor definition) into Methods.

@@

For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
% E: I removed the description of relative extinction results.  We don't discuss it for other models or show it in the main text, and I never really know what to make of that compound parameter.  But it's good that it's in the supp info.

For breeding system alone, we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
Thus, differences in net diversification are best explained by both breeding system and an unknown factor.
% The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
The statistical model comparison shows very strong support for including the hidden state (\cref{table:bayesfactors}).

For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
The association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
%
When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.
Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).

@@

@@

\begin{figure}
\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".  Rosana, if you add the .svg file for this figure, I can make small changes like this.
\caption{Net diversification rates for all models that include diploidization.}  
\label{figure:netdivall}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf} % FIXME: same note as above about "+A/B"
\caption{Net diversification rates for all models that do not include diploidization.}  
\label{figure:netdivnodip}
\end{figure}

@@
